### 2018-02-02 analysis/analysis.Rmd
Add analysis report with data and graphs for promise lifespan
This commit adds a R Markdown report with data and graphs for
promise lifespan analysis.
The promise-lifespan.R script has also been updated.

---
title: "Analyses of Promises in R"
author: "Aviral Goel"
output: 
  html_document:
    theme: flatly
    highlight: tango
    keep_md: true
    toc: true
    toc_float: true
    number_sections: true
params:
  analysis_output_dir: TRUE
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(methods) ## including this prevents a warning - "executing load actions ..."
library(DT)
library(tidyverse)
library(knitr)
```

```{r echo=FALSE, message=FALSE}
# https://www.rdocumentation.org/packages/knitr/versions/1.17/topics/include_graphics
show_graph <- function(analysis_name, graph_filename) {
    file.path(params$analysis_output_dir, analysis_name, "graph", graph_filename) %>%
    include_graphics()
}

get_table <- function(analysis_name, table_filename) {
    file.path(params$analysis_output_dir, analysis_name, "table", table_filename) %>%
    read_csv()
}

show_table <- function(analysis_name, table_filename) {
    get_table(analysis_name, table_filename) %>%
    datatable(rownames = FALSE)
}
```
# Promise Lifespan

This section discusses the lifespan of promises in terms of the number of GC Cycles.
Essentially, we study how long promises are alive in terms of GC cycles, the minimum
number of GC cycles for which they are required to be alive and the number of GC
cycles for which they are alive above and beyond the minimum.

## Categorization

We categorize the promises observed while tracing as follows:

1. **Foreign Promises** are created before the tracing starts. Since we don't see their creation, 
   we can't study their lifespan in terms of GC cycles. They are **excluded from the analysis**
2. **Immortal Promises** are created but not garbage collected while tracing. They are 
   **excluded from the analysis** because we can't compute their lifespan.
3. **Mortal Promises** are created and garbage collected while tracing.
   They are the only promises that are **included in the analysis**.

### Graph

```{r echo=FALSE, message=FALSE, fig.width=10, fig.height=4.5}
show_graph("promise-lifespan", "count.png")
```

**TOTAL PROMISE COUNT** = **FOREIGN PROMISE COUNT** + **IMMORTAL PROMISE COUNT** + **MORTAL PROMISE COUNT**

### Table

#### Summarized

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "counts-summarized.csv")
```

#### Raw

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "counts.csv")
```

### Insights

This categorization sheds light on how many promises are excluded from the analysis.
Clearly, the number of **Immortal Promises** and **Foreign Promises** is quite low.
Any analysis done on **Mortal Promises** should be statistically relevant.

## Alive Promise Count

The following graph shows the number of mortal promises that are alive for a given number of GC cycles.

### Graph

```{r echo=FALSE, message=FALSE, fig.width=10, fig.height=4.5}
show_graph("promise-lifespan", "alive.png")
```

### Table

#### Summarized

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "alive-summarized.csv")
```

#### Raw

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "alive.csv")
```

### Insights

We see that most promises are alive for the shortest possible duration, i.e. - one GC cycle.
The remaining promises are a fraction of the number of promises that are alive for this duration.
We also observe a decreasing trend in the number of promise that are alive for a given number of
GC cycles, as the number of GC cycles increase. This means that most promises are used immediately
and have little need of being alive for longer durations.

## Indispensable Promise Count

The following graph shows the number of mortal promises that need to be alive for a given number of GC cycles.

### Graph

```{r echo=FALSE, message=FALSE, fig.width=10, fig.height=4.5}
show_graph("promise-lifespan", "indispensable.png")
```

### Table

#### Summarized

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "indispensable-summarized.csv")
```

#### Raw

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "indispensable.csv")
```

## Dispensable Promises

The following graph shows the number of promises that extend their required lifespan by a given number of GC cycles.

### Graph

```{r echo=FALSE, message=FALSE, fig.width=10, fig.height=4.5}
show_graph("promise-lifespan", "dispensable.png")
```

### Table

#### Summarized

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "dispensable-summarized.csv")
```

#### Raw

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "dispensable.csv")
```

## Unforced Promises

The following graph shows the number of promises that are never forced.

### Graph

```{r echo=FALSE, message=FALSE, fig.width=10, fig.height=4.5}
show_graph("promise-lifespan", "unforced.png")
```

### Table

#### Summarized

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "unforced-summarized.csv")
```

#### Raw

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "unforced.csv")
```

## Comparison

```{r echo=FALSE, message=FALSE}

indispensable <- 
    get_table("promise-lifespan", "indispensable-summarized.csv") %>%
    rename(`INDISPENSABLE PROMISE COUNT` = `PROMISE COUNT`)

dispensable <- 
    get_table("promise-lifespan", "dispensable-summarized.csv") %>%
    rename(`DISPENSABLE PROMISE COUNT` = `PROMISE COUNT`)

alive <- 
    get_table("promise-lifespan", "alive-summarized.csv") %>%
    rename(`ALIVE PROMISE COUNT` = `PROMISE COUNT`)

alive %>%
full_join(indispensable, by = c("GC CYCLES" = "GC CYCLES")) %>%
full_join(dispensable, by = c("GC CYCLES" = "GC CYCLES")) %>%
datatable(rownames = FALSE)
```