### 2018-02-11 analysis/report/02-memory.Rmd

# Memory

This section compares the number and size of objects allocated by R.

## Object Count Distribution

```{r echo=FALSE, message=FALSE, out.width="575px"}
show_graph("promise-memory-usage", "promise_count.png")
show_graph("promise-memory-usage", "relative_count.png")
```

### Absolute Count Quantiles 
```{r echo=FALSE, message=FALSE}
get_table("promise-memory-usage", "quantiles.csv") %>%
filter(`QUANTITY` == "COUNT") %>%
spread(QUANTILE, VALUE) %>%
select(`OBJECT TYPE`, `0%`, `25%`, `50%`, `75%`, `100%`) %>%
datatable(rownames = FALSE)
```

### Absolute Count Outliers 
```{r echo=FALSE, message=FALSE}
get_table("promise-memory-usage", "outliers.csv") %>%
filter(`QUANTITY` == "COUNT") %>%
select(`OBJECT TYPE`, `SCRIPT`, `COUNT`) %>%
datatable(rownames = FALSE)
```

### Relative Count Quantiles 
```{r echo=FALSE, message=FALSE}
get_table("promise-memory-usage", "quantiles.csv") %>%
filter(`QUANTITY` == "RELATIVE COUNT") %>%
spread(QUANTILE, VALUE) %>%
select(`OBJECT TYPE`, `0%`, `25%`, `50%`, `75%`, `100%`) %>%
datatable(rownames = FALSE)
```

### Relative Count Outliers 
```{r echo=FALSE, message=FALSE}
get_table("promise-memory-usage", "outliers.csv") %>%
filter(`QUANTITY` == "RELATIVE COUNT") %>%
select(`OBJECT TYPE`, `SCRIPT`, `RELATIVE COUNT`) %>%
datatable(rownames = FALSE)
```

## Object Size Distribution

```{r echo=FALSE, message=FALSE, out.width="575px"}
show_graph("promise-memory-usage", "promise_size.png")
show_graph("promise-memory-usage", "relative_size.png")
```

### Absolute Size Quantiles 
```{r echo=FALSE, message=FALSE}
get_table("promise-memory-usage", "quantiles.csv") %>%
filter(`QUANTITY` == "SIZE") %>%
spread(QUANTILE, VALUE) %>%
select(`OBJECT TYPE`, `0%`, `25%`, `50%`, `75%`, `100%`) %>%
datatable(rownames = FALSE)
```

### Absolute Size Outliers 
```{r echo=FALSE, message=FALSE}
get_table("promise-memory-usage", "outliers.csv") %>%
filter(`QUANTITY` == "SIZE") %>%
select(`OBJECT TYPE`, `SCRIPT`, `SIZE`) %>%
datatable(rownames = FALSE)
```

### Relative Size Quantiles 
```{r echo=FALSE, message=FALSE}
get_table("promise-memory-usage", "quantiles.csv") %>%
filter(`QUANTITY` == "RELATIVE SIZE") %>%
spread(QUANTILE, VALUE) %>%
select(`OBJECT TYPE`, `0%`, `25%`, `50%`, `75%`, `100%`) %>%
datatable(rownames = FALSE)
```

### Relative Size Outliers 
```{r echo=FALSE, message=FALSE}
get_table("promise-memory-usage", "outliers.csv") %>%
filter(`QUANTITY` == "RELATIVE SIZE") %>%
select(`OBJECT TYPE`, `SCRIPT`, `RELATIVE SIZE`) %>%
datatable(rownames = FALSE)
```

## Raw Data
```{r echo=FALSE, message=FALSE}
show_table("promise-memory-usage", "object_size_summary.csv")
```