### 2018-03-23 analysis/analysis.Rmd

---
title: "Analyses of Promises in R"
author: "Aviral Goel"
output: 
  html_document:
    theme: flatly
    highlight: tango
    keep_md: true
    toc: true
    toc_float: true
    number_sections: true
params:
  analysis_output_dir: TRUE
---
```{r set-options, echo=FALSE, cache=FALSE}
options(width = 1200)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(methods) ## including this prevents a warning - "executing load actions ..."
library(DT)
library(tidyverse)
library(knitr)
library(png)
library(grid)
library(gridExtra)
```

```{r echo=FALSE, message=FALSE}
# https://www.rdocumentation.org/packages/knitr/versions/1.17/topics/include_graphics

get_graph <- function(analysis_name, graph_filename) {
    file.path(params$analysis_output_dir, analysis_name, "graph", graph_filename)
}

show_graph <- function(analysis_name, graph_filename) {
    get_graph(analysis_name, graph_filename) %>%
    include_graphics()
}

get_table <- function(analysis_name, table_filename) {
    file.path(params$analysis_output_dir, analysis_name, "table", table_filename) %>%
    read_csv()
}

show_table <- function(analysis_name, table_filename) {
    get_table(analysis_name, table_filename) %>%
    datatable(rownames = FALSE)
}
```
# Promise Lifespan

This section discusses the lifespan of promises in terms of the number of GC Cycles.
Essentially, we study how long promises are alive in terms of GC cycles, the minimum
number of GC cycles for which they are required to be alive and the number of GC
cycles for which they are alive above and beyond the minimum.

## Promise Categories

We categorize the promises observed while tracing as follows:

1. **Foreign Promises** are created before the tracing starts. Since we don't see their creation, 
   we can't study their lifespan in terms of GC cycles. They are **excluded from the analysis**
2. **Immortal Promises** are created but not garbage collected while tracing. They are 
   **excluded from the analysis** because we can't compute their lifespan.
3. **Mortal Promises** are created and garbage collected while tracing.
   They are the only promises that are **included in the analysis**.
   They are further classified as follows:
   i. **Forced Promises** are the mortal promises that are looked up at least once.
   ii. **Unforced Promises** are mortal promises that are never looked up.

```{r echo=FALSE, message=FALSE, fig.width=10, fig.height=4.5}
show_graph("promise-lifespan", "counts.png")
```

**TOTAL PROMISE COUNT** = **FOREIGN PROMISE COUNT** + **IMMORTAL PROMISE COUNT** + **MORTAL PROMISE COUNT** + **UNFORCED PROMISE COUNT**

### Summarized Data

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "counts-summarized.csv")
```

### Raw Data

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "counts.csv")
```

## Promise Lifespan Distribution

The following graph shows the lifespan of mortal promises in terms of GC cycles. 
The lifespan of a promise is defined as the number of GC cycles for which the 
promise is alive.

```{r echo=FALSE, message=FALSE, fig.width=10, fig.height=4.5}
show_graph("promise-lifespan", "alive.png")
```

### Summarized Data

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "alive-summarized.csv")
```

### Raw Data

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "alive.csv")
```

## Indispensable Promises

The following graph shows the number of mortal promises that need to be alive for a given number of GC cycles.

```{r echo=FALSE, message=FALSE, fig.width=10, fig.height=4.5}
show_graph("promise-lifespan", "indispensable.png")
```

### Summarized Data

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "indispensable-summarized.csv")
```

### Raw Data

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "indispensable.csv")
```

## Dispensable Promises

The following graph shows the number of promises that extend their required lifespan by a given number of GC cycles.

```{r echo=FALSE, message=FALSE, fig.width=10, fig.height=4.5}
show_graph("promise-lifespan", "dispensable.png")
```

### Summarized Data

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "dispensable-summarized.csv")
```

### Raw Data

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "dispensable.csv")
```

## Unforced Promises

The following graph shows the number of promises that are never forced.

```{r echo=FALSE, message=FALSE, fig.width=10, fig.height=4.5}
show_graph("promise-lifespan", "unforced.png")
```

### Summarized Data

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "unforced-summarized.csv")
```

### Raw Data

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "unforced.csv")
```

## Comparison

```{r echo=FALSE, message=FALSE}

indispensable <- 
    get_table("promise-lifespan", "indispensable-summarized.csv") %>%
    rename(`INDISPENSABLE PROMISE COUNT` = `PROMISE COUNT`)

dispensable <- 
    get_table("promise-lifespan", "dispensable-summarized.csv") %>%
    rename(`DISPENSABLE PROMISE COUNT` = `PROMISE COUNT`)

alive <- 
    get_table("promise-lifespan", "alive-summarized.csv") %>%
    rename(`ALIVE PROMISE COUNT` = `PROMISE COUNT`)

alive %>%
full_join(indispensable, by = c("GC CYCLES" = "GC CYCLES")) %>%
full_join(dispensable, by = c("GC CYCLES" = "GC CYCLES")) %>%
datatable(rownames = FALSE)
```
<!--
# Promise Space Usage

This section compares the number and size of objects allocated by R.

## R Object Count

Count of R objects.

### Graph

```{r echo=FALSE, message=FALSE}
show_graph("promise-memory-usage", "promise-count.png")
```

### Table

#### Summarized

```{r echo=FALSE, message=FALSE}
get_table("promise-memory-usage", "object_size_summary.csv") %>%
select(-`SIZE`, -`SCRIPT`) %>%
group_by(`OBJECT TYPE`) %>%
summarize(`MEAN` = mean(`COUNT`),
          `MEDIAN` = median(`COUNT`),
          `LOWER QUANTILE (25%)` = unname(stats::quantile(sort(`COUNT`), 0.25)),
          `UPPER QUANTILE (75%)` = unname(stats::quantile(sort(`COUNT`), 0.75))) %>%
datatable(rownames=FALSE)
```

#### Raw

```{r echo=FALSE, message=FALSE}
get_table("promise-memory-usage", "object_size_summary.csv") %>%
select(-`SIZE`) %>%
datatable(rownames=FALSE)
```

### Outliers

#### Lower Quantile
```{r echo=FALSE, message=FALSE}
get_table("promise-memory-usage", "lower_count_quantile.csv") %>%
datatable(rownames=FALSE)
```

#### Upper Quantile
```{r echo=FALSE, message=FALSE}
get_table("promise-memory-usage", "lower_count_quantile.csv") %>%
datatable(rownames=FALSE)
```

## R Object Size

Size of R objects.

### Graph

```{r echo=FALSE, message=FALSE}
show_graph("promise-memory-usage", "promise-size.png")
```

### Table

#### Summarized

```{r echo=FALSE, message=FALSE}
get_table("promise-memory-usage", "object_size_summary.csv") %>%
select(-`COUNT`, -`SCRIPT`) %>%
group_by(`OBJECT TYPE`) %>%
summarize(`MEAN` = mean(`SIZE`),
          `MEDIAN` = median(`SIZE`),
          `LOWER QUANTILE (25%)` = unname(stats::quantile(sort(`SIZE`), 0.75)),
          `UPPER QUANTILE (75%)` = unname(stats::quantile(sort(`SIZE`), 0.25))) %>%
datatable(rownames=FALSE)
```

#### Raw

```{r echo=FALSE, message=FALSE}
get_table("promise-memory-usage", "object_size_summary.csv") %>%
select(-`SIZE`) %>%
datatable(rownames=FALSE)
```

### Outliers

#### Lower Quantile
```{r echo=FALSE, message=FALSE}
get_table("promise-memory-usage", "lower_size_quantile.csv") %>%
datatable(rownames=FALSE)
```

#### Upper Quantile
```{r echo=FALSE, message=FALSE}
get_table("promise-memory-usage", "upper_size_quantile.csv") %>%
datatable(rownames=FALSE)
```

## R Object Count

Count of R objects.

### Graph

```{r echo=FALSE, message=FALSE}
show_graph("promise-memory-usage", "promise-count.png")
```

### Table

#### Summarized

```{r echo=FALSE, message=FALSE}
get_table("promise-memory-usage", "object_size_summary.csv") %>%
select(-`SIZE`, -`SCRIPT`) %>%
group_by(`OBJECT TYPE`) %>%
summarize(`MEAN` = mean(`COUNT`),
          `MEDIAN` = median(`COUNT`),
          `LOWER QUANTILE (25%)` = unname(stats::quantile(sort(`COUNT`), 0.25)),
          `UPPER QUANTILE (75%)` = unname(stats::quantile(sort(`COUNT`), 0.75))) %>%
datatable(rownames=FALSE)
```

#### Raw

```{r echo=FALSE, message=FALSE}
get_table("promise-memory-usage", "object_size_summary.csv") %>%
select(-`SIZE`) %>%
datatable(rownames=FALSE)
```

### Outliers

#### Lower Quantile
```{r echo=FALSE, message=FALSE}
get_table("promise-memory-usage", "lower_count_quantile.csv") %>%
datatable(rownames=FALSE)
```

#### Upper Quantile
```{r echo=FALSE, message=FALSE}
get_table("promise-memory-usage", "lower_count_quantile.csv") %>%
datatable(rownames=FALSE)
```

# Environment Function Usage

This section discusses the usage of functions which manipulate environments.

## baseenv()

This is the environment of the base package.

```{r echo=FALSE, message=FALSE, out.width = "400px"}
show_graph("environment", "baseenv.png")
show_graph("environment", "baseenv-relative.png")
```

## emptyenv()

This is the empty environment, the topmost environment.

```{r echo=FALSE, message=FALSE, out.width = "400px"}
show_graph("environment", "emptyenv.png")
show_graph("environment", "emptyenv-relative.png")
```

## new.env(hash = TRUE, parent = parent.frame(), size = 29L)

`new.env` returns a new (empty) environment with (by default) enclosure the parent frame.

```{r echo=FALSE, message=FALSE, out.width = "400px"}
show_graph("environment", "new.env.png")
show_graph("environment", "new.env-relative.png")
```

## NewEnvironment

This is the C level function which gets called when a new environment is created.

```{r echo=FALSE, message=FALSE, out.width = "400px"}
show_graph("environment", "NewEnvironment.png")
show_graph("environment", "NewEnvironment-relative.png")
```

## environment(fun = NULL)

If fun is a function or a formula then `environment(fun)` returns the environment associated with that function or formula. If fun is NULL then the current evaluation environment is returned.

```{r echo=FALSE, message=FALSE, out.width = "400px"}
show_graph("environment", "environment.png")
show_graph("environment", "environment-relative.png")
```

## environment(fun) <- value

`environment(fun) <- value` sets the environment of the function or formula fun to the value given.

```{r echo=FALSE, message=FALSE, out.width = "400px"}
show_graph("environment", "environment<-.png")
show_graph("environment", "environment<--relative.png")
```

## parent.env

`parent.env` returns the enclosing environment of its argument.

```{r echo=FALSE, message=FALSE, out.width = "400px"}
show_graph("environment", "parent.env.png")
show_graph("environment", "parent.env-relative.png")
```

## parent.env(env) <- value

`parent.env<-` sets the enclosing environment of its first argument.
The replacement function parent.env<- is extremely dangerous as it can be used to destructively change environments in ways that violate assumptions made by the internal C code. It may be removed in the near future.

```{r echo=FALSE, message=FALSE, out.width = "400px"}
show_graph("environment", "parent.env<-.png")
show_graph("environment", "parent.env<--relative.png")
```

~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

### Table

#### Summarized

```{r echo=FALSE, message=FALSE}
get_table("environment", "environment-function-usage.csv") %>%
select(-`SCRIPT`) %>%
group_by(`FUNCTION NAME`) %>%
summarize(`MEAN` = mean(`COUNT`),
          `MEDIAN` = median(`COUNT`),
          `LOWER QUANTILE (25%)` = unname(stats::quantile(sort(`COUNT`), 0.25)),
          `UPPER QUANTILE (75%)` = unname(stats::quantile(sort(`COUNT`), 0.75))) %>%
datatable(rownames=FALSE)
```

#### Raw
```{r echo=FALSE, message=FALSE}
show_table("environment", "environment-function-usage-spread.csv")
```

### Outliers

#### Lower Quartile
```{r echo=FALSE, message=FALSE}
get_table("promise-memory-usage", "lower_size_quantile.csv") %>%
datatable(rownames=FALSE)
```

#### Upper Quartile
```{r echo=FALSE, message=FALSE}
get_table("promise-memory-usage", "upper_size_quantile.csv") %>%
datatable(rownames=FALSE)
```
-->