### 2018-03-23 analysis/report/02-memory.Rmd

# Memory

This section compares the number and size of objects allocated by R.

## Object Count Distribution

```{r echo=FALSE, message=FALSE, out.width="575px"}
show_graph("promise-memory-usage", "object_count.png")
show_graph("promise-memory-usage", "relative_object_count.png")
```

## Object Size Distribution

```{r echo=FALSE, message=FALSE, out.width="575px"}
show_graph("promise-memory-usage", "object_size.png")
show_graph("promise-memory-usage", "relative_object_size.png")
```

## Summarized Data
```{r echo=FALSE, message=FALSE}
show_table("promise-memory-usage", "summary.csv")
```

## Raw Data
```{r echo=FALSE, message=FALSE}
show_table("promise-memory-usage", "raw.csv")
```