### 2018-03-23 analysis/report/04-argument-promise-mode.Rmd

# Argument Promise Mode

This section studies the mode of argument promises.
There are two types of argument promise modes:

1. **Default** are the default arguments to functions
2. **Custom** are user defined arguments to functions

## Promise mode count

The graphs below show the number of promises for each mode.
The second graph shows the average number of promises per mode.

```{r echo=FALSE, message=FALSE, out.width="575px"}
show_graph("argument-promise-mode", "mode_total.png")
show_graph("argument-promise-mode", "mode_average.png")
```

### Data (Total)
```{r echo=FALSE, message=FALSE}
get_table("argument-promise-mode", "promise_mode.csv") %>%
group_by(MODE) %>%
summarize(COUNT = sum(COUNT)) %>%
datatable(rownames = FALSE)
```

### Data (Average)

```{r echo=FALSE, message=FALSE}
table <- get_table("argument-promise-mode", "promise_mode.csv")
script_count <- length(unique(table[["SCRIPT"]]))

table %>%
group_by(MODE) %>%
summarize(COUNT = sum(COUNT) / script_count) %>%
datatable(rownames = FALSE)
```

## Promise mode distribution

The graphs below show the distribution of promises across
scripts for each mode. The first one shows the absolute count
and the second one shows the relative count.

```{r echo=FALSE, message=FALSE, out.width="575px"}
show_graph("argument-promise-mode", "mode_distribution.png")
show_graph("argument-promise-mode", "mode_relative_distribution.png")
```

### Data (Absolute)

```{r echo=FALSE, message=FALSE}
get_table("argument-promise-mode", "promise_mode.csv") %>%
group_by(SCRIPT, MODE) %>%
summarize(COUNT = sum(COUNT)) %>%
datatable(rownames = FALSE)
```

### Data (Quantiles)

```{r echo=FALSE, message=FALSE}
get_table("argument-promise-mode", "promise_mode.csv") %>%
group_by(SCRIPT, MODE) %>%
summarize(COUNT = sum(COUNT)) %>%
ungroup() %>%
nest(-`MODE`) %>%
mutate(Quantiles = map(data, ~ quantile(.$COUNT, na.rm = TRUE))) %>%
unnest(map(Quantiles, tidy)) %>%
rename(QUANTILE = names, VALUE = x) %>%
datatable(rownames = FALSE)
```

### Data (Relative)

```{r echo=FALSE, message=FALSE}
get_table("argument-promise-mode", "promise_mode.csv") %>%
group_by(SCRIPT, MODE) %>%
summarize(COUNT = sum(COUNT)) %>%
mutate(COUNT = COUNT/sum(COUNT)) %>%
datatable(rownames = FALSE)
```

## Promise mode type

The graphs below break down the promise modes further into the types
of their expression and value slots.

```{r echo=FALSE, message=FALSE, out.width="575px"}
show_graph("argument-promise-mode", "mode_expression_type.png")
show_graph("argument-promise-mode", "mode_value_type.png")
```

### Data (Expression Type)

```{r echo=FALSE, message=FALSE}
get_table("argument-promise-mode", "promise_mode.csv") %>%
filter(SLOT == "EXPRESSION") %>%
group_by(TYPE, MODE) %>%
summarize(COUNT = sum(COUNT)) %>%
datatable(rownames = FALSE)
```

### Data (Value Type)

```{r echo=FALSE, message=FALSE}
get_table("argument-promise-mode", "promise_mode.csv") %>%
filter(SLOT == "VALUE") %>%
group_by(TYPE, MODE) %>%
summarize(COUNT = sum(COUNT)) %>%
datatable(rownames = FALSE)
```