### 2018-03-23 analysis/report/05-position-evaluation-mode.Rmd

# Position Evaluation Mode

This section discusses the strictness of function parameter
positions.
For each function, strictness of its parameter position is 
computed by observing the evaluation state of all promises
that occupy that position. If the promises at a given
parameter position are always evaluated, then that position
is strict. If they are never evaluated, then that position is
lazy. If they are sometimes evaluated then that position 
is mixed. The graph below summarizes this information across
all functions.

```{r echo=FALSE, message=FALSE}
show_graph("position-evaluation-mode", "evaluation_mode_distribution.png")
```

## Data (Summarized)

```{r echo=FALSE, message=FALSE}
show_table("position-evaluation-mode", "evaluation_mode_summarized.csv")
```

## Data

```{r echo=FALSE, message=FALSE}
show_table("position-evaluation-mode", "evaluation_mode.csv")
```