### 2018-04-03 graphs/report-number-of-accesses-to-a-promise.Rmd

# Number of accesses to a promise

``` {r, "number of accesses to a promise, aggregate", echo=FALSE, warning=FALSE, fig.width=10, fig.height=5}
evaluations <- 
  load_local_csv("promise_evaluations.csv") #%>%
print(typeof(evaluations %>% pull(number)))
print(sum(as.numeric(evaluations %>% pull(number))))

evaluations <- evaluations %>% 
    group_by(no.of.evaluations) %>% 
    summarise(number=sum(as.numeric(number), na.rm=TRUE)) %>% 
    mutate(percent=100*number/sum(number)) %>%
    mutate(labels=no.of.evaluations) %>%
    mutate(no.of.evaluations=as.numeric(as.character(no.of.evaluations)))  %>%
    mutate(no.of.evaluations=ifelse(is.na(no.of.evaluations),
				     max(no.of.evaluations, na.rm=TRUE)+1, 
				                            no.of.evaluations)) %>% 
							    arrange(no.of.evaluations)

ggplot(evaluations, aes(x=no.of.evaluations, y=number)) + 
  geom_col() +
    scale_y_continuous(labels=pp_trunc) +
      scale_x_discrete(limit=0:11, labels=evaluations %>% pull(labels)) +
        theme(axis.title.x=element_blank(), legend.position="none") +
	  ylab("No. of promises") + xlab("No. of accesses")
	  ```

<div style="font-size: 50%">
Promise types follow the type designations in ``Rinternals.h``.  
</div>