### 2018-04-03 graphs/report-promise-return-types.Rmd
Promise return types.

# Promise return types

``` {r, "promise return types, aggregate",  echo=FALSE, fig.width=10, fig.height=5}
return_types <- 
  load_local_csv("return_types.csv")

ggplot(return_types %>%   
         group_by(type) %>% 
         summarise(number=sum(number)) %>% 
         mutate(percent=100*number/sum(number)), 
       aes(x=type, y=number)) + 
  geom_col() +
  scale_y_continuous(labels=pp_trunc) +
  theme(axis.title.x=element_blank(), 
        legend.position="none", 
        axis.text.x=element_text(angle=90, vjust=.5, hjust=1)) +
  ylab("No. of promises") + 
  xlab("Return type")
```