### 2018-04-03 graphs/report-main.Rmd
millions and billions

---
title: "R dyntrace report"
author: "K. Siek, A. Goel"
output: 
  html_document:
    theme: flatly # default cerulean journal flatly readable 
                  # spacelab united cosmo lumen paper sandstone simplex yeti
    highlight: tango
    toc: true
    toc_depth: 3
    number_sections: true
params:
  CSV_DIR: "/home/kondziu/workspace/promise-dyntracing-experiment/graphs/csv/"
  PRINT_DATA: TRUE
---

<style>
  .col2 {
    columns: 2 200px;         /* number of columns and width in pixels*/
    -webkit-columns: 2 200px; /* chrome, safari */
    -moz-columns: 2 200px;    /* firefox */
  }

  div.together {
    display: inline-block;
    width: 100%;
  }
</style>

``` {r, "initial setup", include=FALSE}
#rmarkdown::render("graphs_from_csv.Rmd", params = list(CSV_DIR = "/home/kondziu/workspace/R-dyntrace/reports/aggregate/", PRINT_DATA = TRUE))
#params <- data.frame(PRINT_DATA=TRUE, CSV_DIR="/home/kondziu/workspace/promise-dyntracing-experiment/graphs/csv/")

#knitr::opts_chunk$set(cache=TRUE)

library(dplyr)
library(hashmap)
library(ggplot2)
library(knitr)
library(kableExtra)
library(scales)
library(gridExtra)
library(png)

SEXP_TYPE_NAMES = c(
    "NIL", "SYM", "LIST", "CLOS", "ENV",  "PROM", # 0-5
    "LANG", "SPECIAL", "BUILTIN", "CHAR",  "LGL", # 6-10
    "INT", "REAL", "CPLX", "STR", "DOT", "ANY",   # 13-18
    "VEC", "EXPR", "BCODE", "EXTPTR", "WEAKREF",  # 19-23
    "RAW", "S4")                                  # 24-25

pp <- function(number) format(number, big.mark=",", scientific=FALSE, trim=FALSE, digits=2)
load_local_csv <- function (csv_file) {
    path <- file.path(params$CSV_DIR, csv_file)
    write(paste("reading", path), stderr())
    as_tibble(read.csv(path, stringsAsFactors=FALSE))
}
print_data <- function (x) if(params$PRINT_DATA) print(x, row.names=FALSE)
prettify <- function(x) x %>% mutate(number=pp_trunc(number), percent=pp_perc(percent)) %>% as.data.frame
pp_trunc <- function(x) ifelse(x==0, paste(format(x, digits=2, scientific=FALSE)),
                               ifelse(x < 1000, format(x, digits=2, scientific=FALSE), 
                                      ifelse(x < 1000000, 
                                             paste(format(floor((x/1000)*10)/10, digits=2, scientific=FALSE), "k", sep=""),
                                             ifelse(x < 1000000000,
                                                    paste(format(floor((x/1000000)*10)/10, digits=2, scientific=FALSE), "mil", sep="")))),
                                                    paste(format(floor((x/1000000000)*10)/10, digits=2, scientific=FALSE), "bil", sep="")))),
pp_perc <- function(x) dollar_format(prefix="", suffix="%")(x)
is_outlier <- function(x) x < quantile(x, 0.25) - 1.5 * IQR(x) | x > quantile(x, 0.75) + 1.5 * IQR(x)

#http://mathematicalcoffee.blogspot.cz/2014/06/ggpie-pie-graphs-in-ggplot2.html
ggpie <- function (dat, by, totals, labels=by) {
    ggplot(dat, aes_string(x=factor(1), y=totals, fill=by)) +
        geom_bar(stat='identity', color='black') +
        guides(fill=guide_legend(override.aes=list(colour=NA))) + # removes black borders from legend
        coord_polar(theta='y') +
        theme(axis.ticks=element_blank(),
            axis.text.y=element_blank(),
            axis.text.x=element_text(colour='black'),
            axis.title=element_blank(),
            legend.position='none') +
    scale_y_continuous(breaks=cumsum(dat[[totals]]) - dat[[totals]] / 2, labels=dat[[labels]])
}

make_graphs_for_packages <- function(data, f, max.col=10) {
  graphs <- lapply((data %>% select(name) %>% distinct %>% pull(name)), f(data))
  do.call("grid.arrange", c(graphs, ncol=min(max.col, ceiling(sqrt(length(graphs))))))
}

make_graphs_for_range <- function(data, range_column, f, max.col=10) {
  graphs <- lapply(data %>% select(!!range_column) %>% distinct %>% pull(!!range_column),
                   function(x) f(data %>% filter_(paste(range_column,"==",quote(x)))))
  do.call("grid.arrange", c(graphs, ncol=min(length(graphs), max.col)))
}

# hard-coded columns: number
make_graphs_by_split <- function(data, key, split="name", max_columns=10, graphing_function=NA) {
  if (is.na(graphing_function))
    graphing_function <- 
      function(package, data) 
        ggplot(data, aes_string(x=key, y="number")) + 
        geom_bar(stat="identity") +
        scale_y_continuous(
          labels=pp_trunc, 
          breaks=c(max(data$number))) +
        theme(
          axis.title.y=element_blank(),
          axis.ticks.y=element_blank(),
          axis.text.y=element_text(angle=90, size=5, vjust=.5),
          axis.text.x=element_text(angle=90, size=5, vjust=.5),
          axis.title.x=element_text(size=5),
          legend.position="none") +
        xlab(package)

  packages <- data %>% select_(split) %>% distinct %>% pull(split)
  key_range <- data %>% select_(key) %>% distinct %>% collect

  splitter <- function (package)
    graphing_function(
      package,
      data %>% 
        filter(UQ(rlang::sym(split)) == !!package) %>% 
        group_by_(key) %>%
        summarise(number=sum(number)) %>% 
        right_join(key_range, by=key) %>%
        mutate(number=ifelse(is.na(number), 0, number)))

  graphs <- lapply(packages, splitter)
  do.call("grid.arrange", c(graphs, ncol=min(max_columns, ceiling(sqrt(length(graphs))))))
}

# hard coded: percent
make_graphs_by_range <- function(data, range_column, max_columns=10, graphing_function=NA) {
  if(is.na(graphing_function))
    graphing_function <- function(data)
      ggplot(data, aes_string(range_column, "percent")) + 
      geom_violin() +
      scale_y_continuous(labels=pp_perc) +
      theme(axis.title.x=element_blank(),
            axis.title.y=element_blank(),
            legend.position="none")

  splitter <- function(x) 
    graphing_function(clean_data %>% filter_(paste(range_column, "==", quote(x))))

  range <- data %>% select_(range_column) %>% distinct %>% pull(range_column)
  clean_data <- data %>% mutate(percent = ifelse(is.na(percent), 0, percent))
  graphs <- lapply(range, splitter)
  do.call("grid.arrange", c(graphs, ncol=min(length(graphs), max_columns)))
}
```

<!--```{r child = 'report-basic-info.Rmd', params=params}				      
```-->

```{r child = 'report-promise-accesses.Rmd', params=params}				      
```

```{r child = 'report-number-of-accesses-to-a-promise.Rmd', params=params}				      
```

```{r child = 'report-promise-code-types.Rmd', params=params}				      
```

```{r child = 'report-type-traces.Rmd', params=params}				      
```

```{r child = 'report-promise-return-types.Rmd', params=params}				      
```