### 2018-04-04 graphs/report-interactions-among-promises.Rmd
Reporting etc.

# Interactions among promises

```{r, "interactions among promises", echo=FALSE, comment="", fig.width=10, fig.height=4.5}
# Graph one
promises_forced_by_another_promise <-   
  load_local_csv("promises_forced_by_another_promise.csv", tibble(forced_by_another=logical(), number=numeric(), percent=numeric())) %>% 
  group_by(forced_by_another) %>% 
  summarise(number=sum(as.numeric(number))) %>% 
  mutate(percent=100*number/sum(number))

g_receive <- 
  ggplot(promises_forced_by_another_promise, 
         aes(x=forced_by_another, y=number)) + 
  geom_col() +
  scale_y_continuous(labels=pp_trunc) +
  theme(legend.position="none") +
  ylab("No. of promises") + 
  xlab("This promise forced\nby another promise")

# Graphs two and three
promises_forcing_other_promises <- 
  load_local_csv("promises_forcing_other_promises.csv", tibble(number_of_forced_promises=numeric(), number=numeric(), percent=numeric()))

promises_forcing_other_promises_summarized <- 
  promises_forcing_other_promises %>%
  group_by(number_of_forced_promises) %>% 
  summarise(number=sum(as.numeric(number))) %>% 
  mutate(percent=100*number/sum(number))

# Graph two
promises_forcing_other_promises_small <-
  promises_forcing_other_promises_summarized %>% 
  mutate(number_of_forced_promises=ifelse(number_of_forced_promises>=3, "3+", 
                                          number_of_forced_promises)) %>%
  group_by(number_of_forced_promises) %>%
  summarise(number=sum(as.numeric(number))) %>% 
  mutate(percent=100*number/sum(number))

g_cause_small <- 
  ggplot(promises_forcing_other_promises_small, aes(x=number_of_forced_promises, y=number)) + 
  geom_col() +
  scale_y_continuous(labels=pp_trunc) +
  ylab("No. of promises") + 
  xlab("No. of promises forced\nby this promise") +
  theme(legend.position="none", axis.title.y=element_blank())

# Graph three
promises_forcing_other_promises_big <-
  promises_forcing_other_promises_summarized %>% 
  filter(number_of_forced_promises >= 3) %>%
  group_by(number_of_forced_promises) %>%
  summarise(number=sum(number)) %>% 
  mutate(percent=100*number/sum(number))

g_cause_big <- 
  ggplot(promises_forcing_other_promises_big, aes(x=number_of_forced_promises, y=number)) + 
  geom_col() +
  scale_y_continuous(labels=pp_trunc) +
  ylab("No. of promises") + 
  xlab("No. of promises forced\nby this promise (3+)") +
  theme(legend.position="none", axis.title.y=element_blank())

# Drawing
grid.arrange(g_receive, g_cause_small, g_cause_big, ncol=3, widths=c(4,5,9))

# Printout
promises_forced_by_another_promise %>% prettify %>% kable(col.names=c("forced by another promise","number","percentage"), align="lrr")
promises_forcing_other_promises_summarized %>% prettify %>% kable(col.names=c("promises that force this many promises","number","percentage"), align="lrr")
```

<div style="font-size: 50%">
  Promises forced by other promises:  
  `TRUE` is when a promise was created within some promise x but was forced within a different promise y.   
`FALSE` is when a promise was forced within the same promise as it was created.  
`NA` means the promise was not forced.  
</div>