### 2018-04-04 graphs/report-number-of-accesses-to-a-promise.Rmd

# Number of accesses to a promise

``` {r, "number of accesses to a promise, aggregate", echo=FALSE, warning=FALSE, fig.width=10, fig.height=5}
evaluations <- 
  load_local_csv("promise_evaluations.csv") %>%
  group_by(no.of.evaluations) %>% 
  summarise(number=sum(as.numeric(number), na.rm=TRUE)) %>% 
  mutate(percent=100*number/sum(number)) %>%
  mutate(labels=no.of.evaluations) %>%
  mutate(no.of.evaluations=as.numeric(as.character(no.of.evaluations)))  %>%
  mutate(no.of.evaluations=ifelse(is.na(no.of.evaluations),
                        	  	    max(no.of.evaluations, na.rm=TRUE)+1, 
				                          no.of.evaluations)) %>% 
	arrange(no.of.evaluations)

ggplot(evaluations, aes(x=no.of.evaluations, y=number)) + 
  geom_col() +
  scale_y_continuous(labels=pp_trunc) +
  scale_x_discrete(limit=0:11, labels=evaluations %>% pull(labels)) +
  theme(axis.title.x=element_blank(), legend.position="none") +
	ylab("No. of promises") + xlab("No. of accesses")

evaluations %>% select(labels, number, percent) %>% prettify %>% kable(col.names=c("no. of evaluations","number","percentage"), align="lrr")
```

<div style="font-size: 50%">
Promise types follow the type designations in ``Rinternals.h``.  
</div>