### 2018-04-04 graphs/report-promise-evaluation-distances.Rmd

# Promise evaluation distances

<!--Effective distance is the distance from the call where a promise originated to the call in which it was evaluated. It is measured along the call graph (along inbound edges).  Since the data is gathered from a trace, the call graph is a tree.-->

```{r, "promise actual distances, aggregate", echo=FALSE, fig.width=10, fig.height=4.5}
actual_distances <- 
  load_local_csv("actual_distances.csv", tibble(distance=character(), number=numeric(), percent=numeric())) %>% 
  group_by(actual_distance) %>% 
  summarise(number=sum(as.numeric(number), na.rm=TRUE)) %>% 
  mutate(percent=100*number/sum(number))

ggplot(actual_distances, aes(x=actual_distance, y=number)) + 
  geom_col() +
  scale_x_discrete(limits=c(NA,0:20,">20",-1)) +
  scale_y_continuous(labels=pp_trunc) +
  ylab("No. of promises") + 
  theme(axis.title.x=element_blank())

actual_distances %>% prettify %>% kable(col.names=c("distance","number","percentage"), align="lrr")
```

<div style="font-size: 50%">
  **n** - distance from the function for which a promise was created and the function in which it was evaluated    
**NA** - promise was not executed, there is no distance measure (a virgin promise)  
**-1** - promise was executed but there is no path from the origin to the point of evaluation (an escaped promise)    
</div>