### 2018-04-04 graphs/report-what-do-specific-promises-return.Rmd

# What do specific promises return?

```{r, "what do specific promises return", echo=FALSE, comment="", fig.width=10.5, fig.height=4.5}
types_to_return_types <- 
  load_local_csv("promise_code_to_return_types.csv", tibble(types=character(), number=numeric(), percent=numeric())) %>%
  group_by(types) %>% 
  summarise(number=sum(as.numeric(number))) %>% 
  mutate(percent=100*number/sum(number)) %>%
  arrange(desc(number))

ggplot(types_to_return_types, aes(x=types, y=number)) + 
  geom_col() +
  scale_y_continuous(labels=pp_trunc) +
  theme(axis.title.x=element_blank(), 
        axis.title.y=element_blank(),
        legend.position="none", 
        axis.text.x=element_text(angle=90, vjust=.5, hjust=1)) +
  ylab("No. of promises") + 
  xlab("Type/return type")

types_to_return_types %>% arrange(desc(number)) %>% prettify %>% kable(col.names=c("type/return type","number","percentage"), align="lrr")
```

<div style="font-size: 50%">
  Promise types follow the type designations in ``Rinternals.h``.  
**α→β** indicates that the promise contains an expression of type **α** which evaluates to  an expression of type **β** when the promise is forced.
</div>