### 2018-04-06 graphs/report-call-and-function-definition-strictness.Rmd

# Call and function definition strictness

```{r, "call strictnes", echo=FALSE, fig.width=10, fig.height=4.5}
strictness_labeller <- function(x) 
  filter(x, number>0) %>%
  mutate(label=ifelse(is.na(strict),  
                      ifelse(type=="closure", 
                             "closure\n(no arguments)", 
                             type),  
                      ifelse(strict, 
                             paste0(type, "\n(strict)"),
                             paste0(type, "\n(non-strict)"))))

strict_closure_calls <- 
  load_local_csv("call_strictness_by_type.csv", tibble(strict=logical(), type=character(), number=numeric(), percent=numeric())) %>%
  strictness_labeller

strict_closure_calls_aggregate <- 
  strict_closure_calls %>%
  group_by(label) %>% 
  summarise(number=sum(as.numeric(number))) %>% 
  mutate(percent=100*number/sum(number))

ggplot(strict_closure_calls_aggregate, aes(x=label, y=number)) + 
  ggtitle("Call strictness") +
  geom_col() +
  scale_y_continuous(labels=pp_trunc) +
  theme(legend.position="none", axis.title.x=element_blank()) +
  ylab("No. of calls")

strict_closure_calls_aggregate %>% prettify %>% kable(col.names=c("strictness","number","percentage"), align="lrr")
```

```{r, "function strictnes", echo=FALSE, fig.width=10, fig.height=4.5}
strict_closure_functions <-
  load_local_csv("aggregated_function_strictness_by_type.csv", tibble(strict=logical(), type=character(), number=numeric(), percent=numeric())) %>%
  strictness_labeller

strict_closure_functions_aggregate <-
  strict_closure_functions %>%
  group_by(label) %>%
  summarise(number=sum(as.numeric(number))) %>%
  mutate(percent=100*number/sum(number))

ggplot(strict_closure_functions_aggregate, aes(x=label, y=number)) +
  ggtitle("Function strictness") +
  geom_col() +
  scale_y_continuous(labels=pp_trunc) +
  theme(legend.position="none", axis.title.x=element_blank()) +
  xlab("Function strictness (closures)")

strict_closure_functions_aggregate %>% prettify %>% kable(col.names=c("strictness","number","percentage"), align="lrr")
```

```{r, "function strictnes load", echo=FALSE}
# strict_closure_functions <-
#  load_local_csv("function_strictness_by_type.csv") %>% 
#  strictness_labeller
```

<div style="font-size: 50%">
  A call is **strict** if all of its arguments are **evaluated** and none of the arguments is **escaped**. (This means that a function may pass unevaluated promises to a child call and still be strict, but it cannot leak the promise).  
A function is **strict** if all of its calls are **strict**. Functions that don't take promises are neither **strict** nor **non-strict** -- these can be either arrity-0 closures or built-in (which accept values rather than promises).
</div>

<!--# Call strictness ratio-->

```{r, "strict calls ratio breakdown", echo=FALSE, fig.width=10, fig.height=4.5}
# strict_calls_ratio <- load_local_csv("call_strictness_ratio.csv")
# 
# strict_calls_ratio_split_arg <- strict_calls_ratio %>% 
#   mutate(evaluated_args=lapply(strsplit(strictnesss_ratio, "/"), `[[`, 1),
#          total_args=lapply(strsplit(strictnesss_ratio, "/"), `[[`, 2))
# 
# range <- strict_calls_ratio_split_arg %>% select(total_args) %>% distinct %>% arrange(total_args) %>% top(20)
# strict_calls_
```

## Call strictness ratios

```{r, "strict calls ratio", echo=FALSE, fig.width=10, fig.height=4.5}
strict_calls_ratio <- 
  load_local_csv("call_strictness_ratio.csv", tibble(strictness_ratio=character(), number=numeric(), percent=numeric()))

threshold =
  strict_calls_ratio %>%
  group_by(strictness_ratio) %>%
  summarise(number=sum(as.numeric(number))) %>%
  select(number) %>%
  arrange(desc(number)) %>%
  head(8) %>%
  pull(number) %>%
  min

strict_calls_ratio_labelled <-
  strict_calls_ratio %>%
  group_by(strictness_ratio) %>%
  summarise(number=sum(as.numeric(number)),
  percent=(100*number/sum(number))) %>%
  mutate(short=ifelse(number >= threshold, as.character(strictness_ratio), "other"))

data <- strict_calls_ratio_labelled %>%
  group_by(short) %>%
  summarize(number=sum(as.numeric(number)))

ggplot(data, aes(x=short, y=number)) +
  geom_col() +
  scale_y_continuous(labels=pp_trunc) +
  theme(legend.position="none") +
  ylab("No. of calls") +
  xlab("Ratio of evaluated promises (evaluated args/total args)")

print(data)
#data %>% prettify %>% kable(col.names=c("strictness ratio","number","percentage"), align="lrr")
```

<div style="font-size: 50%">
**Strictness ratio** of the form `n/m` means that a call evaluated `n` promises out of `m` promises that were created for it. `0/0` can mean either that a function had no arguments or that it was a built-in etc (had arguments, but they were passed without promises).
</div>

## Call strictness ratios (top 20 within other)

```{r, "call strictness ratios, other top 20", echo=FALSE, fig.width=10, fig.height=4.5}
#strict_calls_ratio_other_top20 <-
#  strict_calls_ratio_labelled %>%
#  filter(short=="other") %>%
#  arrange(desc(number)) %>%
#  head(20)

# ggplot(strict_calls_ratio_other_top20, aes(x=strictness_ratio, y=number)) +
# geom_col() +
# scale_y_continuous(labels=pp_trunc) +
# theme(legend.position="none") +
# ylab("No. of calls") +
# xlab("Ratio of evaluated promises")
# 
# strict_calls_ratio_other_top20 %>% prettify %>% kable(col.names=c("strictness ratio","number","percentage"), align="lrr")
```

<div style="font-size: 50%">
**Strictness ratio** of the form `n/m` means that a call evaluated `n` promises out of `m` promises that were created for it. `0/0` can mean either that a function had no arguments or that it was a built-in etc (had arguments, but they were passed without promises).
</div>

## Call strictness ratios (all)

```{r, "call strictness ratios, top 20", echo=FALSE, fig.width=10, fig.height=4.5}
# strict_calls_ratio_labelled %>% 
# arrange(desc(number)) %>%
# mutate(number=pp_trunc(number), percent=pp_perc(percent)) %>%
# select(strictness_ratio, number, percent) %>%
# kable(col.names=c( "strictness ratio", "number", "percent"), align="lll")
```

<div style="font-size: 50%">
**Strictness ratio** of the form `n/m` means that a call evaluated `n` promises out of `m` promises that were created for it. `0/0` can mean either that a function had no arguments or that it was a built-in etc (had arguments, but they were passed without promises).
</div>