### 2018-04-10 graphs/report-promise-code-types.Rmd
Reporting saves graphs to PDF.

# Promise code types

``` {r, "promise code types, aggregate", echo=FALSE, fig.width=10, fig.height=5}
promise_types <- 
  load_local_csv("promise_types.csv") %>% 
  group_by(type) %>% summarise(number=sum(as.numeric(number))) %>% mutate(percent=100*number/sum(number))

g <- ggplot(promise_types, aes(x=type, y=number)) + 
  geom_col() +
  scale_y_continuous(labels=pp_trunc) +
  theme(axis.title.x=element_blank(), 
        legend.position="none", 
        axis.text.x=element_text(angle=90, vjust=.5)) +
  ylab("No. of promises") + 
  xlab("Promise type")
g
save_graph("promise-types.pdf", g)

promise_types %>% arrange(desc(number)) %>% prettify %>% kable(col.names=c("promise code type","number","percentage"), align="lrr")
```

<div style="font-size: 50%">
Promise types follow the type designations in ``Rinternals.h``.  
<!--**α→β** indicates that type **α** variously evaluates to type **β**.
**α** indicates that the promise contains the expression of type **α**.  
**→α** indicates that the promise contains a promise which contains the expression of type **α**, etc.
**BCODE →α** indicates that it's a byte-compiled expression that was originally of type **α**.  
**SYM→α** indicates that it's a symbol that points to an expression of type **α**.  -->
</div>