### 2018-04-10 graphs/report-promise-return-types.Rmd

# Promise return types

``` {r, "promise return types, aggregate",  echo=FALSE, fig.width=10, fig.height=5}
return_types <- 
  load_local_csv("return_types.csv", tibble(type=character(), number=numeric(), percent=numeric())) %>%   
  group_by(type) %>% 
  summarise(number=sum(as.numeric(number))) %>% 
  mutate(percent=100*number/sum(number))

g <- ggplot(return_types, 
       aes(x=type, y=number)) + 
  geom_col() +
  scale_y_continuous(labels=pp_trunc) +
  theme(axis.title.x=element_blank(), 
        legend.position="none", 
        axis.text.x=element_text(angle=90, vjust=.5, hjust=1)) +
  ylab("No. of promises") + 
  xlab("Return type")

g
save_graph("promise-return-types.pdf", g)

return_types %>% arrange(desc(number)) %>% prettify %>% kable(col.names=c("promise code type","number","percentage"), align="lrr")
```