### 2018-04-14 old-analysis/report-calls-and-fucntions-by-type.Rmd

# Calls and functions by type

<!--center>
  <div class="col2">
    <div class="together"-->
```{r, "calls by type", echo=FALSE, fig.width=5, fig.height=4}
    calls_types <- load_local_csv("call_types.csv", tibble(type=character(), number=numeric(), percent=numeric()))

    data <- 
      calls_types %>% 
      group_by(type) %>% 
      summarise(number=sum(as.numeric(number))) %>% 
      mutate(percent=100*number/sum(number))

    if (data %>% count %>% pull(n) > 0) {
      g <- ggplot(data, aes(x=type, y=number)) + 
        geom_bar(stat="identity") +
        ylab("No. of calls") + 
        scale_y_continuous(labels=pp_trunc) +
        theme(axis.title.x=element_blank())
      g
      save_graph("call-types.pdf", g)
    }

    if (data %>% count %>% pull(n) > 0) {
      data %>% prettify %>% kable(col.names=c("type","number","percentage"), align="lrr")
    }
```
    <!--/div-->

```{r, "functions by type", echo=FALSE, fig.width=5, fig.height=4}
    function_types <- load_local_csv("aggregated_function_types.csv", tibble(type=character(), number=numeric(), percent=numeric()))

    data <- function_types %>%
      group_by(type) %>% 
      summarise(number=sum(number)) %>% 
      mutate(percent=100*number/sum(number))

    if (data %>% count %>% pull(n) > 0) {
      g <- ggplot(data, aes(x=type, y=number)) + 
        geom_bar(stat="identity") +
        ylab("No. of function definitions") + 
        scale_y_continuous(labels=pp_trunc) +
        theme(axis.title.x=element_blank())
      g
      save_graph("function-types.pdf", g)
    }

    if (data %>% count %>% pull(n) > 0) {
      data %>% prettify %>% kable(col.names=c("type","number","percentage"), align="lrr")
    }
```

      <div style="font-size: 50%">
      **functions** refers to function definitions   
    **calls** refers to function executions/calls
    </div>