### 2018-04-14 old-analysis/report-promise-forces-by-types.Rmd

# Promise forces by types

```{r, "promise forces by types, aggregate",  echo=FALSE, , comment="", fig.width=10, fig.height=5}
force.histogram.by.type <- 
  load_local_csv("forces_by_type.csv", tibble(type=character(), no.of.forces=numeric(), number=numeric(), percent=numeric())) %>%
  group_by(type, no.of.forces) %>% 
  summarise(number=sum(as.numeric(number))) %>% 
  mutate(percent=100*number/sum(number))

g <- ggplot(
  force.histogram.by.type, 
  aes(x=type, y=number, fill=factor(no.of.forces))) + 
  geom_bar(stat="identity", position="dodge") +
  ylab("No. of promises") + 
  scale_fill_discrete(name="No. of forces",
                      breaks=(force.histogram.by.type$no.of.forces %>% unique),
                      labels=(force.histogram.by.type$no.of.forces %>% unique)) +
  scale_y_continuous(labels=pp_trunc) +
  theme(axis.title.x=element_blank(), legend.position = c(0.9, 0.9))
g
save_graph("promise-forces-by-types.pdf", g)
force.histogram.by.type %>% arrange(type, no.of.forces) %>% prettify %>% kable(col.names=c("promise code type", "forces", "number","percentage"), align="lrr")
```

<div style="font-size: 50%">
  Promise types follow the type designations in ``Rinternals.h``.
</div>