### 2018-04-14 old-analysis/report-type-traces.Rmd

# Type traces

```{r, "what's inside promises", echo=FALSE, fig.width=10, fig.height=4.5}
promise_full_types <-
  load_local_csv("promise_full_types.csv") %>%
  group_by(type) %>% 
  summarise(number=sum(as.numeric(number))) %>% 
  mutate(percent=100*number/sum(number)) %>% 
  mutate(type=as.character(type)) #%>%
  #filter(grepl("^PROM", type)) %>%
  #mutate(type=gsub("PROM", "", type))
  #mutate(type=ifelse(len(type) > 15, paste0(strtrim(type, 13), ".."), type))

promise_full_types %>% 
  arrange(desc(number)) %>%
  mutate(number=pp_trunc(number), percent=pp_perc(percent)) %>%
  select(number, percent, type) %>%
  kable(col.names=c("number", "percent", "type trace"), align="lll")
```

<div style="font-size: 50%">
Promise types follow the type designations in ``Rinternals.h``.  
<!--**α→β** indicates that type **α** variously evaluates to type **β**.  -->
**α** indicates that the promise contains the expression of type **α**.  
**PROM→α** indicates that the promise contains a promise which contains the expression of type **α**, etc.  
**BCODE →α** indicates that it's a byte-compiled expression that was originally of type **α**.    
**SYM→α** indicates that it's a symbol that points to an expression of type **α**.  
**...** is a type that could not has not been correctly determined.
</div>