### 2018-10-08 report/report.Rmd
Add report with function-viewer application

---
title: "Laziness in R"
output:
  html_document:
    toc: true
    theme: united
params:
  output_directory: ../latest
  image_format: png
editor_options:
  chunk_output_type: console
---

<style type="text/css">
.main-container {
  max-width: 1200px;
  margin-left: auto;
  margin-right: auto;
  padding-left: 15px;
  padding-right: 15px;
}

.shiny-frame {
    height: 550px;
}
</style>

```{r echo = FALSE, message = FALSE}

library(DT)
library(fs)
library(png)
library(stringr)
library(dplyr)
library(magrittr)
library(rlang)
library(knitr)

```

```{r paths, echo = FALSE}

function_definitions_dirpath <- path(params$output_directory, "function-definitions")
parameter_usage_dirpath <- path(params$output_directory, "parameter-usage")
visualizations <- path(parameter_usage_dirpath, "visualizations")
summary <- path(parameter_usage_dirpath, "summary")

```

```{r datatsets, echo = FALSE}
import_data_table <- function(name) {
    write(path(summary, name), "log.txt")
    promisedyntracer::read_data_table(path(summary, name))
}

closure_force_order_distribution <-
     import_data_table("closure_force_order_distribution.csv") %>%
     mutate(force_order = str_replace_all(str_sub(force_order, 2), "\\|", " ~ "))

closure_force_order_distribution_summary <-
     import_data_table("closure_force_order_distribution_summary.csv")

closures <- import_data_table("closures.csv")

closure_call_count_summary_distribution <- 
    import_data_table("closure_call_count_summary_distribution.csv") %>%
    mutate(relative_function_count = round(relative_function_count * 100, 3))

closure_argument_count_summary_distribution <- 
    import_data_table("closure_argument_count_summary_distribution.csv") %>%
    mutate(relative_function_count = round(relative_function_count * 100, 3))

classification_summary <- import_data_table("classification_summary.csv")

parameter_mode_distribution_summary <-
    import_data_table("parameter_mode_distribution_summary.csv")

argument_mode_distribution_summary <-
    import_data_table("argument_mode_distribution_summary.csv")
#

# closure_classification_summary_distribution <- import_data_table("closure_classification_summary_distribution.csv")
# 
# position_classification_summary_distribution <- import_data_table("position_classification_summary_distribution.csv")
# 
# summary <- import_data_table("summary.csv")
#

parameter_force_order_summary <- 
     import_data_table("closure_force_order_distribution.csv") %>%
     mutate(force_order = str_replace_all(str_sub(force_order, 2), "\\|", " ~ "))

function_class_summary_distribution <-
    import_data_table("function_class_summary_distribution.csv") %>%
    mutate(relative_function_count = round(relative_function_count * 100, 3))

parameter_class_summary_distribution <-
    import_data_table("parameter_class_summary_distribution.csv") %>%
    mutate(relative_parameter_count = round(relative_parameter_count * 100, 3))

argument_class_distribution_summary <-
    import_data_table("argument_class_distribution_summary.csv")

argument_use_mode_distribution_summary <- 
     import_data_table("argument_use_mode_distribution_summary.csv")
```

```{r echo=FALSE}
count_labels <-
  Vectorize(function(x, digits = 2) {

    paste_round <- function(value, div, suffix, sep)
        paste(round(value/div, digits), suffix, sep = sep)

    if(is.na(x)) {
      "NA"
    } else if(x < 10^3) {
      paste0(x)
    } else if(x < 10^6) {
      paste_round(x, 1000, "K", sep=" ")
    } else if(x < 10^9) {
      paste_round(x, 10^6, "M", sep=" ")
    } else {
      paste_round(x, 10^9, "B", sep=" ")
    }
  },
  "x")
```

This report summarizes the findings on the use of laziness in R.

Data from 23 packages and 41 vignettes.

* Number of functions: `r count_labels(length(unique(closures$function_id)))` 
* Number of calls: `r count_labels(sum(closures$call_count))`
* Number of parameters: `r count_labels(sum(closures$parameter_count))`
* Number of arguments: `r count_labels(sum(closures$parameter_count * closures$call_count))`

## Closures

The table below gives the relevant information for all these functions.

```{r closure_data, echo = FALSE}
datatable(closure_force_order_distribution) #, width="1000px")
```

## Closure call count distribution

The following figures show the distribution of the number of calls made to closures.

```{r echo = FALSE, out.width = "1000px"}
include_graphics(path(visualizations, "closure_call_distribution.png"))
```

```{r echo = FALSE}
datatable(closure_call_count_summary_distribution)
```

### Observations

* Very few functions, **`r closure_call_count_summary_distribution %>% filter(call_count == 1) %>% pull(relative_function_count) %>% first()`%** , are called only once. We can remove these functions from the remaining analyses without affecting the representativity of our results.

* Majority of functions, **`r closure_call_count_summary_distribution %>% filter(!(call_count %in% c("1", "> 10"))) %>% pull(relative_function_count) %>% sum()`%**, are called more than once.

* **`r closure_call_count_summary_distribution %>% filter(call_count == "> 10") %>% pull(relative_function_count) %>% sum()`%** functions are called more than 10 times.

## Closure parameter count distribution

The following figures show the distribution of the number of parameters of closures.

```{r echo = FALSE, out.width = "1000px"}
include_graphics(path(visualizations, "closure_parameter_count_distribution.png"))
```

```{r echo = FALSE}
datatable(closure_argument_count_summary_distribution)
```

### Observations
- Majority of functions, **`r closure_argument_count_summary_distribution %>% filter(parameter_count %in% c(1, 2, 3)) %>% pull(relative_function_count) %>% sum()`%**, have one, two or three parameters.
- A very small number of functions, **`r closure_argument_count_summary_distribution %>% filter(parameter_count == 0) %>% pull(relative_function_count) %>% first()`%**, have zero parameters.

## Closure force order distribution

```{r echo = FALSE, out.width = "1000px"}
include_graphics(path(visualizations, "closure_force_order_distribution.png"))
```

```{r echo = FALSE}
datatable(closure_force_order_distribution_summary)
```

<!--## Parameter Mode

```{r echo = FALSE, out.width = "1000px"}
include_graphics(path(visualizations, "parameter_mode_distribution.png"))
```

```{r echo = FALSE}
datatable(parameter_mode_distribution_summary)
```
-->
## Argument Mode

```{r echo = FALSE, out.width = "1000px"}
include_graphics(path(visualizations, "argument_mode_distribution.png"))
```

```{r echo = FALSE}
datatable(argument_mode_distribution_summary)
```

## Argument Class

```{r echo = FALSE, out.width = "1000px"}
include_graphics(path(visualizations, "argument_class_distribution.png"))
```

```{r echo = FALSE}
datatable(argument_class_distribution_summary)
```

## Argument Use Mode

```{r echo = FALSE, out.width = "1000px"}
include_graphics(path(visualizations, "argument_use_mode_distribution.png"))
```

```{r echo = FALSE}
datatable(argument_use_mode_distribution_summary)
```

## Parameter Class

This section discusses function parameter position usage. The parameter of a function in R can be:
1. looked up for its value
2. looked up for the argument expression
3. not looked up

The graph below shows how the parameter positions are distributed across these categories in all calls to functions.

```{r echo=FALSE}
# datatable(classification_summary)
```

```{r echo = FALSE, out.width = "1000px"}
include_graphics(path(visualizations, "parameter_class_distribution.png"))
```

```{r echo = FALSE}
datatable(parameter_class_summary_distribution)
```

## Argument Class

```{r echo = FALSE, out.width = "1000px"}
include_graphics(path(visualizations, "argument_count_by_mode_class_and_use.png"))
```

```{r echo = FALSE}
datatable(argument_use_mode_distribution_summary)
```

## Function Class

```{r echo = FALSE, out.width = "1000px"}
include_graphics(path(visualizations, "function_class_distribution.png"))
```

```{r echo = FALSE, out.width = "1000px"}
include_graphics(path(visualizations, "function_class_call_distribution.png"))
```

```{r echo = FALSE}
datatable(function_class_summary_distribution)
```