### 2019-03-09 report/escaped_arguments.Rmd

---
title: "The Promise that R makes: Escaped Arguments"
output:
  html_document:
    toc: true
    theme: united
editor_options:
  chunk_output_type: console
---

<style type="text/css">
.main-container {
  max-width: 1200px;
  margin-left: auto;
  margin-right: auto;
  padding-left: 15px;
  padding-right: 15px;
}

.datatables {
    height: 575px !important; 
}
</style>

```{r echo = FALSE, out.width="1000px"}
show_graph("escaped_argument_function_call_count_by_return_value_type")
```

```{r echo = FALSE}
show_table("escaped_argument_function_call_count_by_return_value_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("escaped_argument_function_count_by_return_value_type")
```

```{r echo = FALSE}
show_table("escaped_argument_function_count_by_return_value_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("escaped_argument_function_count_by_category")
```

```{r echo = FALSE}
show_table("escaped_argument_function_count_by_category")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("escaped_argument_count_by_nature")
```

```{r echo = FALSE}
show_table("escaped_argument_count_by_nature")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("escaped_argument_count_by_dispatch_type")
```

```{r echo = FALSE}
show_table("escaped_argument_count_by_dispatch_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("escaped_argument_count_by_expression_type")
```

```{r echo = FALSE}
show_table("escaped_argument_count_by_expression_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("escaped_argument_count_by_value_type")
```

```{r echo = FALSE}
show_table("escaped_argument_count_by_value_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("escaped_argument_count_by_force_type")
```

```{r echo = FALSE}
show_table("escaped_argument_count_by_force_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("escaped_argument_count_by_metaprogram_type")
```

```{r echo = FALSE}
show_table("escaped_argument_count_by_metaprogram_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("escaped_argument_count_by_lookup_type")
```

```{r echo = FALSE}
show_table("escaped_argument_count_by_lookup_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("escaped_argument_count_by_direct_self_scope_mutation_type")
```

```{r echo = FALSE}
show_table("escaped_argument_count_by_direct_self_scope_mutation_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("escaped_argument_count_by_indirect_self_scope_mutation_type")
```

```{r echo = FALSE}
show_table("escaped_argument_count_by_indirect_self_scope_mutation_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("escaped_argument_count_by_direct_lexical_scope_mutation_type")
```

```{r echo = FALSE}
show_table("escaped_argument_count_by_direct_lexical_scope_mutation_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("escaped_argument_count_by_indirect_lexical_scope_mutation_type")
```

```{r echo = FALSE}
show_table("escaped_argument_count_by_indirect_lexical_scope_mutation_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("escaped_argument_count_by_direct_non_lexical_scope_mutation_type")
```

```{r echo = FALSE}
show_table("escaped_argument_count_by_direct_non_lexical_scope_mutation_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("escaped_argument_count_by_indirect_non_lexical_scope_mutation_type")
```

```{r echo = FALSE}
show_table("escaped_argument_count_by_indirect_non_lexical_scope_mutation_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("escaped_argument_count_by_direct_self_scope_observation_type")
```

```{r echo = FALSE}
show_table("escaped_argument_count_by_direct_self_scope_observation_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("escaped_argument_count_by_indirect_self_scope_observation_type")
```

```{r echo = FALSE}
show_table("escaped_argument_count_by_indirect_self_scope_observation_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("escaped_argument_count_by_direct_lexical_scope_observation_type")
```

```{r echo = FALSE}
show_table("escaped_argument_count_by_direct_lexical_scope_observation_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("escaped_argument_count_by_indirect_lexical_scope_observation_type")
```

```{r echo = FALSE}
show_table("escaped_argument_count_by_indirect_lexical_scope_observation_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("escaped_argument_count_by_direct_non_lexical_scope_observation_type")
```

```{r echo = FALSE}
show_table("escaped_argument_count_by_direct_non_lexical_scope_observation_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("escaped_argument_count_by_indirect_non_lexical_scope_observation_type")
```

```{r echo = FALSE}
show_table("escaped_argument_count_by_indirect_non_lexical_scope_observation_type")
```