### 2019-03-17 report/summary.Rmd
Add summary report, fix parameters graphs and add a functions graph

---
title: "The Promise that R makes: Summary"
output:
  html_document:
    toc: true
    theme: united
editor_options:
  chunk_output_type: console
---

<style type="text/css">
.main-container {
  max-width: 1200px;
  margin-left: auto;
  margin-right: auto;
  padding-left: 15px;
  padding-right: 15px;
}

.datatables {
    height: 575px !important; 
}
</style>

## Objects

```{r echo = FALSE, out.width="1000px"}
show_graph("object_count_by_type")
```

```{r echo = FALSE}
show_table("object_count_by_type")
```

## Functions

```{r echo = FALSE, out.width="1000px"}
show_graph("function_call_count_by_type")
```

```{r echo = FALSE}
show_table("function_call_count_by_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("function_count_by_type")
```

```{r echo = FALSE}
show_table("function_count_by_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("function_call_count_by_method_type")
```

```{r echo = FALSE}
show_table("function_call_count_by_method_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("function_count_by_method_type")
```

```{r echo = FALSE}
show_table("function_count_by_method_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("closure_call_count_by_return_value_type")
```

```{r echo = FALSE}
show_table("closure_call_count_by_return_value_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("closure_count_by_call_count")
```

```{r echo = FALSE}
show_table("closure_count_by_call_count")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("closure_count_by_formal_parameter_count")
```

```{r echo = FALSE}
show_table("closure_count_by_formal_parameter_count")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("closure_count_by_return_value_type")
```

```{r echo = FALSE}
show_table("function_count_by_return_value_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("closure_count_by_wrapper_and_force_order")
```

```{r echo = FALSE}
show_table("closure_count_by_wrapper_and_force_order")
```

## Promises

```{r echo = FALSE, out.width="1000px"}
show_graph("promise_count_by_category")
```

```{r echo = FALSE}
show_table("promise_count_by_category")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("argument_promise_count_by_expression_type")
```

```{r echo = FALSE}
show_table("argument_promise_count_by_expression_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("argument_promise_count_by_value_type")
```

```{r echo = FALSE}
show_table("argument_promise_count_by_value_type")
```

```{r echo = FALSE}
show_table("argument_promise_count_by_call_depth")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("argument_promise_count_by_promise_depth")
```

```{r echo = FALSE}
show_table("argument_promise_count_by_promise_depth")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("promise_count_by_force_count")
```

```{r echo = FALSE}
show_table("promise_count_by_force_count")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("promise_count_by_metaprogram_count")
```

```{r echo = FALSE}
show_table("promise_count_by_metaprogram_count")
```
## Arguments

```{r echo = FALSE, out.width="1000px"}
show_graph("argument_count_by_type")
```

```{r echo = FALSE}
show_table("argument_count_by_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("non_missing_argument_count_by_dot_dot_dot")
```

```{r echo = FALSE}
show_table("non_missing_argument_count_by_dot_dot_dot")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("promise_argument_count_by_nature")
```

```{r echo = FALSE}
show_table("promise_argument_count_by_nature")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("promise_argument_count_by_sharing")
```

```{r echo = FALSE}
show_table("promise_argument_count_by_sharing")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("promise_argument_count_by_force_type")
```

```{r echo = FALSE}
show_table("promise_argument_count_by_force_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("promise_argument_count_by_dispatch_type")
```

```{r echo = FALSE}
show_table("promise_argument_count_by_dispatch_type")
```

## Parameters

```{r echo = FALSE, out.width="1000px"}
show_graph("argument_count_by_usage")
```

```{r echo = FALSE}
show_table("argument_count_by_usage")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("formal_parameter_count_by_usage")
```

```{r echo = FALSE}
show_table("formal_parameter_count_by_usage")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("formal_parameter_count_by_usage_class")
```

```{r echo = FALSE}
show_table("formal_parameter_count_by_usage_class")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("closure_count_distribution_by_usage_class")
```

```{r echo = FALSE}
show_table("closure_count_distribution_by_usage_class")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("argument_execution_time_distribution_by_parameter_lookup_class")
```

```{r echo = FALSE}
show_table("execution_times")
```

## Escaped Arguments

```{r echo = FALSE, out.width="1000px"}
show_graph("escaped_argument_function_call_count_by_return_value_type")
```

```{r echo = FALSE}
show_table("escaped_argument_function_call_count_by_return_value_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("escaped_argument_function_count_by_return_value_type")
```

```{r echo = FALSE}
show_table("escaped_argument_function_count_by_return_value_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("escaped_argument_count_by_nature")
```

```{r echo = FALSE}
show_table("escaped_argument_count_by_nature")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("escaped_argument_count_by_force_type")
```

```{r echo = FALSE}
show_table("escaped_argument_count_by_force_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("escaped_argument_count_by_metaprogram_type")
```

```{r echo = FALSE}
show_table("escaped_argument_count_by_metaprogram_type")
```

```{r echo = FALSE, out.width="1000px"}
show_graph("escaped_argument_count_by_lookup_type")
```

```{r echo = FALSE}
show_table("escaped_argument_count_by_lookup_type")
```