### 2019-08-06 report/alien-promises.Rmd
Add context sensitive lookup, lifecycle analysis

---
title: "On the Design, Implementation and Use of Laziness in R"
output:
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
    collapsed: true
    smooth_scroll: false
    number_sections: true
params:
    summarized_data_dirpath: ""
    visualized_data_dirpath: ""
    latex_macro_filepath: ""
---

```{r results = "asis", echo = FALSE}
cat("
<style>
div.section {
  text-align: justify;
  text-justify: inter-word;
}
</style>
")
```

```{r echo = TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(fig.width = 12, fig.height = 4)

library(tidyverse)
library(ggplot2)

PETA <- 10^15
TERA <- 10^12
GIGA <- 10^9
MEGA <- 10^6
KILO <- 10^3
HECTO <- 10^2
DECA <- 10^1
DECI <- 10^-1
CENTI <- 10^-2
MILLI <- 10^-3
MICRO <- 10^-6
NANO <- 10^-9
PICO <- 10^-12
SECONDS_PER_MINUTE <- 60
MINUTES_PER_HOUR <- 60
HOURS_PER_DAY <- 24
SECONDS_PER_HOUR <- SECONDS_PER_MINUTE * MINUTES_PER_HOUR
SECONDS_PER_DAY <- HOURS_PER_DAY * SECONDS_PER_HOUR

paste_round <- function(value, div, digits, suffix, sep = " ") {
    paste(round(value/div, digits), suffix, sep = sep)
}

memory_size_labels <-  function(x, digits = 1) {
    ifelse(x == 0, "0 B",
           { y <- log2(x)
             units <- c("B", "KB", "MB", "GB", "TB", "PB")
             unit <- units[y %/% 10 + 1]
             value <- 2 ^ (y %% 10)
             paste(round(value, digits), unit, sep = " ") })
  }

count_labels <- Vectorize(function(x, digits = 2) 
    if(is.na(x))       "NA"
    else if(x < 10^3)  paste0(x)
    else if(x < 10^6)  paste_round(x, 1000, digits, "K")
    else if(x < 10^9)  paste_round(x, 10^6, digits, "M")
    else if(x < 10^12) paste_round(x, 10^9, digits, "B")
    else               paste_round(x, 10^12, digits, "T"),
  "x")

relative_labels <- function(x) percent_labels(x * 100)
percent_labels <-  function(x) paste0(x, "%", sep="")

time_labels <- Vectorize(function(x, digits = 1)
   if (is.na(x))                   "NA"
   else if (x < NANO)              paste_round(x, PICO, digits, "ps")
   else if (x < MICRO)             paste_round(x, NANO, digits, "ns")
   else if (x < MILLI)             paste_round(x, MICRO, digits, "μs")
   else if(x < 1)                  paste_round(x, MILLI, digits, "ms")
   else if(x < SECONDS_PER_MINUTE) paste_round(x, 1, digits, "s")
   else if(x < SECONDS_PER_HOUR)   paste_round(x, SECONDS_PER_MINUTE, digits, "mins")
   else if(x < SECONDS_PER_DAY)    paste_round(x, SECONDS_PER_HOUR, digits, "hours")
   else                            paste_round(x, SECONDS_PER_DAY, digits, "days"),
  "x")

format_as_percentage <- function(x, digits = 1) {
    rounded <- round(x, digits = digits)
    first_decimal_digit = as.integer(rounded * 10) %% 10
    str_c(format(100 * x,
                 digits = digits,
                 nsmall = first_decimal_digit != 0,
                 scientific = FALSE), "%")
}

format_as_absolute <- function(x, digits = 1) count_labels(x, digits)

format_as_size <- function(x, digits = 1) memory_size_labels(x, digits)

read_table <- function(filename, binary = TRUE, compression_level = 3) {
    read_data_table(path(params$summarized_data_dirpath, filename),
                    binary = binary, 
                    compression_level = compression_level)
}

save_local_graph <- function(plot, name, device = "pdf", ...) {
    ggsave(filename = file.path(params$visualized_data_dirpath, name),
           plot = plot, device = device, ...)
    plot
}

show_table <- function(tablename) {
    read_csv(tablename, col_names = TRUE) %>%
    datatable(fillContainer = TRUE, filter = "top")
}

show_graph <- function(graphname) include_graphics(graphname)

show_two_graphs <- function(first_graphname, second_graphname) {
    load_graph <- function(graphname) {
        grid::rasterGrob(as.raster(readPNG(graphname)), interpolate = TRUE)
    }
    grid.arrange(load_graph(first_graphname),
                 load_graph(second_graphname),
                 ncol = 2)
}

old_theme <- theme_set(theme_bw() + theme(text = element_text(size = 15), panel.border = element_blank()))

# Latex Macro Generation
## name prefix is one of:
## - Perc - adds a % at the end
## - Count - adds labels like K, M, B for Kilo, Million and Billion
## - Size - adds lables like KB, MB, GB for memory sizes
generate_macro <- function(macro_name, value, subst = TRUE) {
    if(subst) {
        name <- as.character(substitute(macro_name))
    } else {
        name <- macro_name
    }
    if (str_sub(name, 1, 4) == "Perc")      formatted_value <- format_as_percentage(value)
    else if(str_sub(name, 1, 5) == "Count") formatted_value <- format_as_absolute(value)
    else if(str_sub(name, 1, 4) == "Size")  formatted_value <- format_as_size(value)
    else stop(str_c("Unidentified name", name, sep = " "))
    formatted_value <- str_replace(formatted_value, "\\%", "\\\\%")
    definition <- str_glue("\\newcommand {{\\{name}}} {{{formatted_value}\\xspace}} %% {value}")
    write_lines(definition, params$latex_macro_filepath, append = TRUE)
}

summarize_outliers <- function(df, grouping_column,
                               count_column, relative_count_column,
                               outlier_value) {

    grouping_column <- enquo(grouping_column)
    count_column <- enquo(count_column)
    relative_count_column <- enquo(relative_count_column)

    outlier_row <-
        df %>%
        filter(!! grouping_column > outlier_value) %>%
        mutate(!!count_column := as.numeric(!!count_column)) %>%
        summarize(!!grouping_column := str_c(">", toString(outlier_value), sep = " "),
                  !!count_column := sum(!!count_column),
                  !!relative_count_column := sum(!!relative_count_column))

    df <-
        df %>%
        mutate(!! grouping_column := as.character(!! grouping_column)) %>%
        bind_rows(outlier_row)

    df
}

bar <- function(widths, total, labels) {
    options(device = cairo_pdf)
    ## If you want to draw arbitrary rectangles, use geom_tile() or geom_rect()
    xmin <- cumsum(c(0, head(widths, -1)))
    df <- data.frame(
        xmin = xmin,
        xmax = cumsum(widths),
        ymin = -0.5,
        ymax = 0.5,
        fill = factor(1:length(widths)),
        label = labels,
        xmid = widths/2.0 + xmin,
        ymid = 0
    )

    plot <-
        df %>%
        ggplot() +
        geom_rect(aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = fill)) +
        geom_text(aes(label = label, x = xmid, y = ymid), colour = "white") +
        scale_y_continuous() +
        scale_x_continuous(breaks = cumsum(widths),
                           labels = relative_labels,
                           sec.axis = sec_axis(trans = ~ . * total,
                                               breaks = cumsum(widths),
                                               labels = count_labels)) +
        theme(axis.text.y = element_blank(),
              axis.ticks = element_blank(),
              panel.grid.minor = element_blank(),
              panel.grid.major = element_blank(),
              legend.position = "none") +
        labs(x = "", y = "")
plot
}

```

```{r echo = TRUE, warning=FALSE, message=FALSE}
alien_promises_by_expression <- read_table("alien_promises_by_expression")
```

# Alien Promises

```{r, eval = TRUE, echo = FALSE}
alien_promises_by_expression %>%
datatable()
```